Verification of Stencil Computations through Spatial Specifications

### 2016-07-07 paper.tex
eval edits

We assigned the specification inferred above to the generated Jacobi
stencils. CamFort checked all 6,561 programs against this spec and found that only 24 of them were
valid implementations of the stencil specification. This was the
expected result as there are $4!$ possible correct permutations
of the four terms in the kernel. All of the other programs were found
to violate the specification in some way. Another way to test the
programs is through visual examination of their output in graphical
format. Figure~\ref{fig:jacobi} shows some of the results of running
the Jacobi iteration programs, rendered into image files. All of the
correct variants emitted images indistinguishable from
Figure~\ref{fig:jacobicorrect}, while all of the incorrect variants
had wildly different image outputs. Both the ``visual test'' and a
byte-by-byte comparison showed results that completely concurred with
the results of the CamFort test.

\begin{figure}[t]
  \centering
  %\subcaptionbox{\label{fig:jacobicorrect}}{\includegraphics[width=0.1\textwidth]{jacobi-correct.png}}
  %\subcaptionbox{}{\includegraphics[width=0.1\textwidth]{jacobi-incorrect1.png}}
  %\subcaptionbox{}{\includegraphics[width=0.1\textwidth]{jacobi-incorrect2.png}}
  %\subcaptionbox{}{\includegraphics[width=0.1\textwidth]{jacobi-incorrect3.png}}
  \subfigure[]{\label{fig:jacobicorrect}\includegraphics[width=0.1\textwidth]{jacobi-correct.png}}
  \subfigure[]{\includegraphics[width=0.1\textwidth]{jacobi-incorrect1.png}}
  \subfigure[]{\includegraphics[width=0.1\textwidth]{jacobi-incorrect2.png}}
  \subfigure[]{\includegraphics[width=0.1\textwidth]{jacobi-incorrect3.png}}
  \caption{Greyscale output of several perturbations of a Jacobi
    iteration program. The correct output is shown in
    \subref{fig:jacobicorrect}.}
  \label{fig:jacobi}
\end{figure}

@@

The Unified Model contains an implementation of the Smagorinsky subgrid-scale model for calculating turbulence. The source file in question yields 39 stencils from 340 lines of code. We selected one as an example:

\begin{minted}[breakindent=0em,linenos=false,xleftmargin=0em]{fortran}
!=stencil readOnce,(forward(depth=1,dim=3,irreflexive))*(backward(depth=1,dim=1))*(backward(depth=1,dim=2))+(forward(depth=1,dim=3))*(backward(depth=1,dim=1))*(reflexive(dim=2))+(forward(depth=1,dim=3))*(backward(depth=1,dim=1,irreflexive))*(backward(depth=1,dim=2,irreflexive))+(forward(depth=1,dim=3))*(backward(depth=1,dim=2))*(reflexive(dim=1))+(backward(depth=1,dim=1))*(backward(depth=1,dim=2,irreflexive))*(reflexive(dim=3))+(backward(depth=1,dim=1,irreflexive))*(backward(depth=1,dim=2))*(reflexive(dim=3))::x
\end{minted}

This stencil specifies the access pattern to a 3-dimensional array (which we have renamed to be called \texttt{x}). The array is accessed 8 times in the computation. We tried introducing indexing errors by changing the offset in one dimension of one of the references to \texttt{x}. For example changing \mintinline{fortran}{x(i+1,j,k)} to \mintinline{fortran}{x(i,j,k)} or \mintinline{fortran}{x(i-1,j,k)}. There are 48 possible errors of this type of which we found only 6 were not detected by the stencil check.
% /um/trunk/src/atmosphere/diffusion_and_filtering/turb_smagorinsky.F90
% Tried introducing indexing errors replacing one of the indices of cx_rho_instance with -1,0,+1
% 42 cases were caught
% 6 cases were not caught

>>>>

We assigned the specification inferred above to the generated Jacobi
stencils programs. CamFort checked all 6,561 programs against this specification and found that only 24 of them were
valid implementations of the stencil specification. This was the
expected result as there are $4!$ possible correct permutations
of the four terms in the kernel. All of the other programs were found
to violate the specification in some way. Another way to test the
programs is through visual examination of their output in graphical
format. Figure~\ref{fig:jacobi} shows some of the results of running
the Jacobi iteration programs, rendered into image files. All of the
correct variants emitted images indistinguishable from
Figure~\ref{fig:jacobicorrect}, while all of the incorrect variants
had wildly different image outputs. Both the ``visual test'' and a
byte-by-byte comparison showed results that completely concurred with
the results of the CamFort test.

\begin{figure}[t]
  \centering
  %\subcaptionbox{\label{fig:jacobicorrect}}{\includegraphics[width=0.1\textwidth]{jacobi-correct.png}}
  %\subcaptionbox{}{\includegraphics[width=0.1\textwidth]{jacobi-incorrect1.png}}
  %\subcaptionbox{}{\includegraphics[width=0.1\textwidth]{jacobi-incorrect2.png}}
  %\subcaptionbox{}{\includegraphics[width=0.1\textwidth]{jacobi-incorrect3.png}}
  \subfigure[]{\label{fig:jacobicorrect}\includegraphics[width=0.1\textwidth]{jacobi-correct.png}}
  \subfigure[]{\includegraphics[width=0.1\textwidth]{jacobi-incorrect1.png}}
  \subfigure[]{\includegraphics[width=0.1\textwidth]{jacobi-incorrect2.png}}
  \subfigure[]{\includegraphics[width=0.1\textwidth]{jacobi-incorrect3.png}}
  \caption{Greyscale output of several perturbations of a Jacobi
    iteration program. The correct output is shown in
    \subref{fig:jacobicorrect}.}
  \label{fig:jacobi}
  \vspace{-0.7em}
\end{figure}

@@

The Unified Model has an implementation of the Smagorinsky
subgrid-scale model for calculating turbulence on which our
inference yields 39 specs from 340 lines of code. One example is:

\begin{minted}[breakindent=0em,linenos=false,xleftmargin=0em,breakafter=)]{fortran}
!=stencil readOnce,(forward(depth=1,dim=3,irreflexive))*(backward(depth=1,dim=1))*(backward(depth=1,dim=2))+(forward(depth=1,dim=3))*(backward(depth=1,dim=1))*(reflexive(dim=2))+(forward(depth=1,dim=3))*(backward(depth=1,dim=1,irreflexive))*(backward(depth=1,dim=2,irreflexive))+(forward(depth=1,dim=3))*(backward(depth=1,dim=2))*(reflexive(dim=1))+(backward(depth=1,dim=1))*(backward(depth=1,dim=2,irreflexive))*(reflexive(dim=3))+(backward(depth=1,dim=1,irreflexive))*(backward(depth=1,dim=2))*(reflexive(dim=3))::x
\end{minted}
%
This specifies the access pattern to a 3-dimensional array (which we have renamed to be called \texttt{x}). The array is accessed 8 times in the computation. We tried introducing indexing errors by changing the offset in one dimension of one of the references to \texttt{x}. For example changing \mintinline{fortran}{x(i+1,j,k)} to \mintinline{fortran}{x(i,j,k)} or \mintinline{fortran}{x(i-1,j,k)}. There are 48 possible errors of this type of which we found only 6 were not detected by the specification checking.
% /um/trunk/src/atmosphere/diffusion_and_filtering/turb_smagorinsky.F90
% Tried introducing indexing errors replacing one of the indices of cx_rho_instance with -1,0,+1
% 42 cases were caught
% 6 cases were not caught
